Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-09-19 Paper.qmd
Added new plots and added householdID to anthropometrics for use in the GAMs.

In order to elicit more support, children often use signals. Signals are a key source of information for individuals in social species and are commonly used to indicate one's value as a mate, one’s formidability to competitors, or one’s worthiness of investment to parents and other social partners [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a]. Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, whereas cues are pieces of information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995]. Crying is likely a signal of need[\[1\]](#_ftn1), for instance, while bleeding would be a cue of need.

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables. While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory. The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.

@@

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for household. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.

@@

# Check: SD (Observations) in the table
```{r}
#| label: tbl-longitudinal
#| tbl-cap: Regression tables for the models examining height and weight growth.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  HeightMean_2024 = "2024 height",
  HeightMean_2023 = "2023 height",
  WeightMean_KG_2024 = "2024 weight",
  WeightMean_KG_2023 = "2023 weight"
)

>>>>

In order to elicit more support, children often use signals. Signals are a key source of information for individuals in social species and are commonly used to indicate one's value as a mate, one’s formidability to competitors, or one’s worthiness of investment to parents and other social partners [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a]. Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, whereas cues are pieces of information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995]. Crying is likely a signal of need[^1], for instance, while bleeding would be a cue of need.

[^1]: We expect “signals of need” to actually be signals of resource valuation. However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables. While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory [@tukey1977]. The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.

@@

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for household [@zeileis2002]. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.

@@

# Check: SD (Observations) in the table

```{r}
#| label: tbl-longitudinal
#| tbl-cap: Regression tables for the models examining height and weight growth.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  HeightMean_2024 = "2024 height",
  HeightMean_2023 = "2023 height",
  WeightMean_KG_2024 = "2024 weight",
  WeightMean_KG_2023 = "2023 weight"
)
